---
title: "Homework 2"
author: "Christopher Crowe"
date: "September 25, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

The piping in the following code chunk:

* reads in the data
* cleans the variable names
* chooses a selection of variables of interest
* converts the entry variable from character to logical

```{r subway data entry}
subway_data = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", col_types = "cccddcccccccccccccccccccccccddcc") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>%
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

As a result of the above code chunk, the data frame `subway_data` contains the following variables: `line`, `station_name`, `station_latitude`, `station_longitude`, `route1`, `route2`, `route3`, `route4`, `route5`, `route6`, `route7`, `route8`, `route9`, `route10`, `route11`, `entry`, `vending`, `entrance_type`, and `ada`. The data has been cleaned using the `janitor::clean_names()` function, which transforms all variable names to a format that is more preferable for data science purposes (e.g., `station_name` instead of `Station Name`). At this point, the data frame `subway_data` has the following dimensions: rows = `r nrow(subway_data)` and columns = `r ncol(subway_data)`. While the data have been somewhat cleaned, they are not tidy because we have the route variable spanning multiple columns.

Using this data, the following code can be used to gather the following information:

```{r distinct, eval = FALSE}
subway_data %>% 
  distinct(line, station_name) %>% 
  nrow()
```

* There are `r subway_data %>% distinct(line, station_name) %>% nrow()` distinct stations.

```{r ada, eval = FALSE}
subway_data %>%
  select(line, station_name, ada) %>% 
  filter(ada == TRUE) %>% 
  distinct(line, station_name) %>% 
  nrow()
```

* `r subway_data %>% select(line, station_name, ada) %>% filter(ada == TRUE) %>% distinct(line, station_name) %>% nrow()` stations are ADA compliant.


```{r vending, eval = FALSE}
subway_data %>% 
  select(line, station_name, vending, entry) %>% 
  filter(vending == "NO") %>% 
  summarize(round(mean(entry), digits = 2))
```

* `r subway_data %>% select(line, station_name, vending, entry) %>% filter(vending == "NO") %>% summarize(round(mean(entry), digits = 2))` of all entrances/exits without vending allow entrance.

The following code chunks reformat the data frame into a more tidy format.

```{r format}
formatted_subway_data = subway_data %>% 
  gather(key = route_number, value = route_name, route1:route11)
```

```{r A train, eval = FALSE}
formatted_subway_data %>% 
  select(line, station_name, route_name) %>% 
  filter(route_name == "A") %>% 
  distinct(line, station_name) %>% 
  nrow()
```

* There are `r formatted_subway_data %>% select(line, station_name, route_name) %>% filter(route_name == "A") %>% distinct(line, station_name) %>% nrow()` distinct stations that serve the A train.

```{r A train ADA, eval = FALSE}
formatted_subway_data %>% 
  select(line, station_name, route_name, ada) %>% 
  filter(route_name == "A" & ada == TRUE) %>% 
  distinct(line, station_name) %>% 
  nrow()
```

* Of these stations, `r formatted_subway_data %>% select(line, station_name, route_name, ada) %>% filter(route_name == "A" & ada == TRUE) %>% distinct(line, station_name) %>% nrow()` are ADA compliant.


## Problem 2

The piping in the following code chunk:

* reads in the Mr. Trash Wheel sheet
* omits columns with notes
* cleans the variable names
* omits rows without dumpster-specific data
* rounds the number of sports balls to the nearest integer and converts it to an interger variable

```{r trash data entry}
trash_wheel = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", range = "Mr. Trash Wheel!A2:N258") %>% 
  janitor::clean_names()
 
```

